Optimal Peak Shaving with Bifacial Solar PV: Simulated Storage Battery Dispatch for Electric Vehicle Charging Load at the Jet Propulsion Laboratory

### 2023-12-20 Paper/paper.tex
paper v1 sent to ema

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{geometry}

\author{}
\date{}

@@

\hypertarget{abstract}{%
  \section{Abstract}\label{abstract}}

Electric load peak shaving is a challenging and important problem
considering the rapid growth of high power electric vehicle charging
sites. Co-located solar photovoltaic provides locally produced energy,
but an additional storage battery is required for reliable peak shaving.
With occasional evening load peaks and an electric tariff with a
16:00-21:00 peak time of use period, vertical bifacial modules facing
West have an economic advantage due to more of the production shifted
into the late afternoon and early evening. Exploiting this, an optimal
electric load peak shaving simulator dispatches a stationary battery to
minimize the retail electric cost incurred by a real electric vehicle
charging site over 10 months. Five different case studies of co-located
bifacial solar photovoltaic arrays are simulated, a base case oriented
South and tilted at 20° and four experimental cases where a percentage
(25\%, 50\%, 75\%, or 100\%) of the array is oriented West and tilted at
90°. The problem is formulated such that the optimizer minimizes the
retail electric cost by adjusting three peak power thresholds, each
corresponding to a time of use period with an associated power price in
\(\$/kW\). The optimizer is implemented using a Newton-Raphson gradient
descent approach and the optimality of the method is verified to within
0.1 \(kW\). A sensitivity analysis on battery energy capacity results in
a largest cost reduction, relative to the South 20° base case, of \$1422
(6.7\%) achieved by the 50\% West 90° case and a modest 25 kWh battery.
A second sensitivity analysis on solar power capacity shows that the
cost reduction increases monotonically and logarithmically for all cases
and battery capacities. The largest cost decrease is \$2579 (16.1\%),
achieved by the 75\% West 90° case at 200\% of nominal solar capacity
along with a 125 kWh battery. Both data and code are shared to the
public on GitHub.

\hypertarget{introduction}{%
  \section{Introduction}\label{introduction}}

\hypertarget{why-peak-load}{%
  \subsection{Why peak load}\label{why-peak-load}}

The renewable energy transition will oversee a global shift toward
electric energy consumption and renewable electric energy production in
the coming decades. Since most electric transmission and distribution
networks were previously built slowly over the span of many decades,
electric load growth will almost definitely outpace network upgrades.
And where these upgrades are completed they are necessarily an
additional cost paid by all electricity consumers, due to the decrease
in the load factor. Furthermore in some markets where companies can own
both generation and distribution, business-as-usual infrastructure
upgrades may be prioritized over the new construction of more complex
and financially risky renewable generators. So while there are some
creative solutions around dynamic capacity limits, two important
solution areas for overall deployment speed, cost effectiveness, and
overall decarbonization goals are (a) more local energy production and
(b) increasing the load factor.

Electric vehicle (EV) charging and electric heating and cooling,
including traditional air conditioning and heat pumps, are two common
new loads that will strain networks in most countries. Heat pump and air
conditioning load factor may be increased with architectural features
such as insulation and thermal storage, but building retrofits are often
slowed down due to permitting and other challenges. Meanwhile extreme
weather events, especially heat waves, will likely further reduce the
load factor of these devices. Many EV users will choose to slowly charge
overnight at home for convenience and to minimize their energy cost.
However workplace, fleet, and public EV charging stations will likely
still be required, and these may especially suffer from a low load
factor due to the convenience or need of fast charging at high power.

\hypertarget{peak-shaving}{%
  \subsection{Peak shaving}\label{peak-shaving}}

Peak load management, or peak shaving, essentially requires choosing a
power threshold and holding the load power below it. Controllable loads,
energy storage, or generation assets behind the billing meter can all be
used to reduce the load power to the threshold power. The threshold may
only apply for only certain time periods. There may be multiple
thresholds and periods each day, month, or year. There are two general
cases of peak shaving worth considering. The most generation formulation
of peak shaving control is formulated in Equation \ref{eg:peakshaving-general}.

\begin{equation}
  \label{eg:peakshaving-general}
  \begin{split}
    I_{threshold,t} &> I_{j,t} - \Sigma_i I_{gen,i,t} - \Sigma_j I_{gen,up,j,t} - \Sigma_k I_{load,dn,k,t}  \\
    \\
    &\text{where:} \\
    &i\text{ is generation asset with no flexibility} \\
    &j\text{ is generation asset with upward flexibility} \\
    &k\text{ is controllable load asset with downward flexibility} \\
    &t\text{ is relevant timesteps}
  \end{split}
\end{equation}

Here we consider the case of no controllable load, a single battery,
solar which reduces the site load, and all values in units of average
real power over the interval \(\Delta h = 1\ hour\). This is descrbied in Equation \ref{eq:peakshaving}.

\begin{equation}
  \label{eq:peakshaving}
  \begin{split}
    P_{threshold,h} &> P_{load,h} - P_{solar,h} - P_{batt,discharge,h} + P_{batt,charge,h}   \\
    &\text{where:} \\
    P_{batt,discharge,h} &\ge 0 \\
    P_{batt,charge,h} &\le 0 \\
    h &\in \{0,1,2,...23\} \\
  \end{split}
\end{equation}

\hypertarget{technical}{%
  \subsubsection{Technical}\label{technical}}

Technical peak shaving refers to the case where a load must operate
under a technical limitation such as a maximum power agreement or
distribution transformer size. The load power must remain under the
threshold at all times, otherwise there may be a technical failure such
an activated overcurrent protection. Even if the load is technically
able to rise above the threshold, doing so may violate a contract
regarding maximum load power. The important consideration is that the
economic cost of failure to hold the load under the threshold is
prohibitively high. The time resolution of technical peak shaving
control and modeling may need to be as low as seconds or milliseconds.
Although this may be a challenging problem if the current limit is
dynamically set or if a larger network is considered, from the
perspective of dispatching the assets to shave the peak the problem is a
relatively simple one: economic dispatch such that the the load current
remains below the threshold current. Technical peak shaving might be
performed on current or apparent power rather than active power.

\hypertarget{economic}{%
  \subsubsection{Economic}\label{economic}}

Rather, economic peak shaving aims to reduce what a consumer pays for
power and possibly also energy. Medium and large electric consumers
often pay a price on energy (€\(/kWh\)) and a price on power
(€\(/kWh_{peak}\)), also called a demand charge. The energy cost
(€\(/kWh \times E_{consumed}\)) may vary with time of day, day of week,
and season of the year, which is often referred to as time of use or
peak pricing. Where there is a sufficient spread between the peak and
off-peak prices there may be the opportunity to curtail load during high
prices, use controllable loads to shift from a high price period to a
lower one, or to use energy storage to buy energy at the lower price and
reduce load during a higher price period. A peak shaving approach
applied to the energy cost could be effective and maybe even
advantageous. However there are several key differences between an
energy based approach and a power one, where peak shaving is better
suited for the latter.

Instead, the power cost (€\(/kW \times P_{max}\) ) typically applies to
the max power during the billing period, where the peak power is the
maximum non-moving average in a given period (e.g. 12:00-18:00 on
weekdays) calculated on a given interval (e.g. 60 minutes). Similar to
the energy cost, there may be multiple time of use periods and
associated prices, such as peak, mid-peak, and off-peak. And where the
spread price is sufficiently high, the period peak can be reduced with
load curtailment or rescheduling, distributed generation such as solar,
or energy storage.

\begin{table}
    \centering
    \caption{Economic and Technical Peak Shaving Comparison}
    \label{tab:econ-tech-peak-shaving}
    \begin{tabularx}{\textwidth}{XXX}
    \toprule
    Feature & Technical Peak Shaving & Economic Peak Shaving  \\
    \midrule
    Consequence of exceeding the threshold & Damage to infrastructure,possible outage & Monetary cost depending on tariff \\
    Averaging interval                     & 1 minute                                 & 15 minutes or 1 hour (typical) \\
    Valid times of day                     & All                                      & Limited (e.g. 16:00 to 21:00) \\
    Peak is reset every..                  & Never                                    & Month, year, day (typical) \\
    \bottomrule
    \end{tabularx}
\end{table}

\hypertarget{methodology}{%
  \section{Methodology}\label{methodology}}

\begin{figure}
  \centering
  \includegraphics[width=5cm,height=\textheight]{./images/oneline.png}
  \caption{Charging Site One Line Schematic. Simplified
  one-line schematic of the simulated EV charging station with true
  measured EV charging power, modelled on-site solar generation, and a
  modelled stationary battery for economic peak shaving.}
  \label{fig:oneline}
\end{figure}

The methodology of this study is a set of offline simulations involving
(a) EV charging power measurements, (b) modelled solar photovoltaic (PV)
production power, and (b) AC-coupled battery dispatch, all behind the
retail electric meter. These are represented in in Figure \ref{fig:oneline}.
In each simulation the battery power is chosen
such that the net load (natural load less solar production, Equation \ref{eq:net-load}) is held
below a given threshold, which is optimally chosen by a gradient descent
optimizer.

\begin{equation}
  \label{eq:net-load}
  \begin{split}
    NL &= L - S \\
    &\text{where:} \\
    &L(t) \text{ is EV charging load time series (non-negative)}\ [kW]\\
    &S(t) \text{ is solar production time series}\ [kW] \\
  \end{split}
\end{equation}

The optimal threshold changes for each time of use (TOU)
period. A single simulation is defined for one solar configuration and
battery capacity and lasts one calendar month, which is the shortest
period for which the cost function is defined. Simulations are repeated
for several months of data, and many different solar and battery design
sizes, allowing for a retail energy cost comparison among different
design scenarios. EV charging power or solar are never curtailed. The
primary decision variables in each time step are (a) battery charge or
discharge, and (b) battery power.

\hypertarget{data}{%
  \subsection{Data}\label{data}}

\hypertarget{ev-charging-load}{%
  \subsubsection{\texorpdfstring{EV Charging Load
    }{EV Charging Load }}\label{ev-charging-load}}

The measured EV charging power ("load") is from the Caltech Adaptive
Charging Network database at the Jet Propulsion Lab (JPL) site. Each
charging session provides time series active power, averaged over a 10
second interval. In the cases when complete time series data is not
available for a session, the charging profile is estimated and the total
charging energy delivered is the same. The disaggregated session time
series data is summed into a single time series of total site charging
power, which is then averaged over 15-minute intervals. Time stamp
indices refer to the beginning of the 15-minute interval. The data
period is from 2018-5-1 00:00 to 2019-2-28 23:45.

\hypertarget{solar-production}{%
  \subsubsection{Solar Production}\label{solar-production}}

>>>>

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[journal,article,submit,pdftex,moreauthors]{Definitions/mdpi}

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arm, arthropoda, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, cyber, dairy, data, ddc, dentistry, dermato, dermatopathology, designs, devices, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, future, futureinternet, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, grasses, gucdd, hazardousmatters, healthcare, hearts, hemato, hematolrep, heritage, higheredu, highthroughput, histories, horticulturae, hospitals, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijpb, ijtm, ijtpp, ime, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jcto, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd, jzbg, kidneydial, kinasesphosphatases, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing,\gdef\@continuouspages{yes}} nanomaterials, ncrna, ndt, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, %%nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, %oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology,\gdef\@ISSN{2813-0618}\gdef\@continuous pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, platforms, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, receptors, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, sclerosis, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, spectroscj, sports, standards, stats, std, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, targets, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal. Changing "submit" to "accept" before posting will remove line numbers.

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.

%=================================================================
% MDPI internal commands - do not modify
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2023}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{ } 
\daterevised{ } % Comment out if no revised date
\dateaccepted{ } 
\datepublished{ } 
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
% cleveref: load \crefname definitions after \begin{document}

\usepackage[acronym]{glossaries}
\usepackage{hhline}
\usepackage{threeparttable} % to use table notes
%\usepackage[draft]{changes} % usage: \deleted{old} \added{new} \replaced{new}{old} also \hl{} for highlighted text
%\usepackage[final]{changes}

%=================================================================
% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Optimal Peak Shaving with Bifacial Solar PV: Simulated Storage Battery Dispatch for Electric Vehicle Charging Load at the Jet Propulsion Laboratory}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Optimal Peak Shaving With Storage Battery and Bifacial Solar PV}

\newcommand{\orcidauthorA}{0000-0003-0237-4148} % Add \orcidA{} behind the author's name
\newcommand{\orcidauthorB}{0000-0002-2106-0374} % Add \orcidB{} behind the author's name

\newcommand{\orcidauthorC}{0000-0002-7883-0034}

% Authors, for the paper (add full first names)
\Author{
  Michael Wood $^{1,2,*}$\orcidA{},
  Emanuele Ogliari $^{1}$\orcidB{},
  Travis Simpkins $^{2}$, and
  Sonia Leva $^{1}$\orcidC{}}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Michael Wood, Emanuele Ogliari, Travis Simpkins, and Sonia Leva}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Wood, M.; Ogliari, E.; Simpkins, T.; Leva, S.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Energy, Politecnico di Milano, Via Lambruschini 4a, 20156, Milan, Italy; emanuelegiovanni.ogliari@polimi.it (E.O.); sonia.leva@polimi.it (S.L.) \\
$^{2}$ \quad muGrid Analyics LLC., 14143 Denver West Parkway Ste 100, Golden, CO, USA; travis@mugrid.com}

% Contact information of the corresponding author
\corres{Correspondence: michael.wood@polimi.it}

\abstract{Electric load peak shaving is a challenging and important problem considering the rapid growth of high-power electric vehicle charging sites. A co-located solar array provides locally produced energy, but an additional storage battery is required for reliable peak shaving. With occasional evening load peaks and an electric tariff with a 16:00-21:00 peak time of use period, vertical bifacial modules facing West have an economic advantage due to more of the production being shifted into the late afternoon and early evening. Exploiting this, an optimal electric load peak shaving simulator dispatches a stationary battery to minimize the retail electric cost incurred by a real electric vehicle charging site over 10 months. Five different case studies of co-located bifacial solar photovoltaic arrays are simulated, a base case oriented South and tilted at 20° and four experimental cases where a percentage (25\%, 50\%, 75\%, or 100\%) of the array is oriented West and tilted at 90°. The problem is formulated such that the optimizer minimizes the retail electric cost by adjusting three peak power thresholds, each corresponding to a time of use period with an associated power price in \(\$/kW\). The optimizer is implemented using a Newton-Raphson gradient descent approach and the optimality of the method is verified to within 0.1 \(kW\). A sensitivity analysis on battery energy capacity results in the largest cost reduction, relative to the South 20° base case, of \$1422 (6.7\%) achieved by the 50\% West 90° case and a modest 25 kWh battery. A second sensitivity analysis on solar power capacity shows that the cost reduction increases monotonically and logarithmically for all cases and battery capacities. The largest cost decrease is \$2579 (16.1\%), achieved by the 75\% West 90° case at 200\% of nominal solar capacity along with a 125 kWh battery. Both data and code are shared with the public on GitHub.}

% Keywords
\keyword{electric load; peak shaving; bifacial solar PV; EV charging;}

@@

\section{Introduction}\label{introduction}

%\subsection{Why peak load}\label{why-peak-load}

The renewable energy transition will oversee a global shift toward electric energy consumption and renewable electric energy production in the coming decades. Since most electric transmission and distribution networks were built slowly over many decades, electric load growth will almost definitely outpace network upgrades. And where these upgrades are completed they are necessarily an additional cost paid by all electricity consumers, due to the decrease in the load factor. Furthermore, in some markets where companies can own both generation and distribution, business-as-usual infrastructure upgrades with cost plus or rate of return regulation may be prioritized over the new construction of more complex and financially riskier renewable generators \cite{Wagner2019}. For these reasons, two important solution areas for overall clean generation deployment speed and cost-effectiveness are distributed energy production and peak load reduction, or peak shaving.

Two new categories of load growth that will strain electric networks in most countries are electric heating and cooling, including traditional air conditioning and heat pumps, and Electric vehicle (EV) charging \cite{Bobmann2015}. Heat pump and air conditioning load factor may be increased with architectural features such as insulation and thermal storage, but building efficiency retrofits often experience barriers such as regulation and financing mechanisms due to permitting and other challenges \cite{Bertone2016}. Meanwhile, extreme weather events, especially heat waves, will likely further reduce the load factor of these loads \cite{Villa2022}. EV charging load differs in that consumers will have the flexibility to slowly charge overnight at home for convenience and to minimize their energy costs. However, workplace, fleet, and public EV charging stations will likely still be required and will experience large peaks as consumers require the convenience of fast charging at stations capable of hundreds of kW per vehicle.

%\hypertarget{peak-shaving}{%
%\subsection{Peak shaving}\label{peak-shaving}%}

Peak load management, or peak shaving, is the process of reducing the maximum load at a given point in an electrical network. This will necessarily increase the load factor,$P_{load,avg} P_{load,max}^{-1}$, even if load curtailment is the only method used. Other methods include load scheduling, load shifting using energy storage, and reduction of the effective load with distributed generation (DG) \cite{Uddin2018}. Peak shaving may include the concept of a peak load threshold which defines a maximum desired peak load during a period, typically on the order of hours. Grid operators may use peak shaving to increase power quality, defer future infrastructure upgrades, decrease losses, or reduce energy costs. End-users typically reduce their peaks solely on a cost basis. Lastly, some carbon reduction is possible by reducing the on/off transitions of peaker plants due to load peaks \cite{Uddin2018}. Broadly speaking there are two categories of peak shaving goals: technical and economic.

%\hypertarget{technical}{%
%\subsubsection{Technical}\label{technical}%}

Technical peak shaving refers to the case where a load must operate under a technical limitation such as a maximum power agreement or distribution transformer size. The load power must remain under the threshold at all times, otherwise, overcurrent protection may be activated or permanent damage to equipment such as lines or transformers. Even if the current is technically able to rise above the threshold, doing so may violate a contract regarding maximum load power. The important consideration is that the economic cost of exceeding the threshold is prohibitively high. The time resolution of technical peak shaving control may need to be as low as seconds or milliseconds \cite{Iannarelli2020}. Although technical peak shaving is complicated by a large network, many assets, or a dynamic threshold, from the perspective of dispatching the assets to shave the peak the problem is relatively simple. Technical peak shaving might be performed on current or apparent power rather than active power.

\begin{table}
  \centering
  \caption{Economic and Technical Peak Shaving Comparison}
  \label{tab:econ-tech-peak-shaving}
  \begin{tabularx}{\textwidth}{XXX}
    \toprule
    Feature                                & Technical Peak Shaving                   & Economic Peak Shaving             \\
    \midrule
    Consequence of exceeding the threshold & Damage to infrastructure, possible outage & Monetary cost depending on tariff \\
    Time resolution                     & $\mu s$ to 10 s                                & 15-60 min or 1 h (typical)    \\
    Valid times of day                     & All                                      & Limited (e.g. 16:00 to 21:00)     \\
    Peak is reset every..                  & Never                                    & Month, year, day (typical)        \\
    \bottomrule
  \end{tabularx}
\end{table}

%\hypertarget{economic}{%
%\subsubsection{Economic}\label{economic}%}

Economic peak shaving aims to reduce what a consumer pays for power and or energy. Medium and large electric consumers often pay an energy price ($p_e\ [\$/kWh]$) and a power price ($p_p\ [\$/kW]$), also called a demand charge. Each may vary with time of day, day of week, and season of the year, which is often referred to as time of use (TOU) or peak pricing. The power cost ($p_p max(P_{load}\ [\$])$) typically applies to the max power during the billing period, where the max power is the maximum non-moving average calculated on a given interval, such as 60 minutes. Where there is a sufficient spread between the peak and off-peak prices there may be the opportunity to load reschedule, shift, or curtail, or dispatch DG.

This work considers a common end-user, economic peak shaving application seen in North America which is described in Equation \ref{eq:peakshaving}. The load may not be curtailed or rescheduled, but solar DG naturally reduces the load and a storage battery is dispatched in a threshold-based peak shaving strategy. The minimum time interval of the threshold is 1 hour.

\begin{equation}
  \label{eq:peakshaving}
  \begin{split}
    P_{threshold,h} & \ge P_{load,h} - P_{solar,h} - P_{batt,discharge,h} + P_{batt,charge,h}                     \\
                    & \text{where:}                                                                               \\
    P               & \text{ is average power (threshold, load, solar, battery) }\ [kW]                             \\
    0               & \le  P_{load,h}, P_{solar,h}, P_{batt,discharge,h}, P_{batt,discharge,h}, P_{batt,charge,h} \\
    h               & \in \{0,1,2,...,23\}                                                                        \\
  \end{split}
\end{equation}

%\hypertarget{methodology}{%
\section{Methodology}\label{methodology}%}

\begin{figure}
  \centering
  \includegraphics[width=7.5cm]{./images/oneline.png}
  \caption{Charging Site One Line Schematic. Simplified one-line schematic of the simulated EV charging station with true measured EV charging power, modeled on-site solar generation, and a modeled stationary battery for economic peak shaving.}
  \label{fig:oneline}
\end{figure}

The methodology of this study is a set of offline simulations involving (a) EV charging power measurements, (b) modeled solar photovoltaic (PV) production power, and (b) AC-coupled battery dispatch, all behind a retail electric meter. These are represented in Figure \ref{fig:oneline}. In each simulation, the battery power is chosen such that the net load (true load less solar production, Equation \ref{eq:net-load}) is held below a given threshold, which is optimally chosen by a gradient descent optimizer.

\begin{equation}
  \label{eq:net-load}
  \begin{split}
    NL & = L - S                                                           \\
       & \text{where:}                                                     \\
       & NL(t) \text{ is the net load time series (non-negative)}\ [kW] \\
       & L(t) \text{ is the EV charging load time series (non-negative)}\ [kW] \\
       & S(t) \text{ is the solar production time series}\ [kW]                \\
  \end{split}
\end{equation}

The optimal threshold changes for each TOU period. A single simulation is defined for one solar configuration and battery capacity and lasts one calendar month, which is the shortest period for which the cost function is defined. Simulations are repeated for several months of data, and many different solar and battery design sizes, allowing for a retail energy cost comparison among different design scenarios. EV charging power and solar production are never curtailed. The primary decision variable in each time step is battery charge or discharge power, which is not necessarily formulated as a binary variable. Nonetheless, the battery can only charge, discharge, or do nothing in each time step.

%\hypertarget{data}{%
\subsection{Data}\label{data}%}

%\hypertarget{ev-charging-load}{%
%\subsubsection{\texorpdfstring{EV Charging Load
%    }{EV Charging Load }}\label{ev-charging-load}%}
\subsubsection{{EV Charging Load }}\label{ev-charging-load}%}

The measured EV charging power ("load") is from the Caltech Adaptive Charging Network database at the Jet Propulsion Lab (JPL) site \cite{Lee2021}. Each charging session provides time series active power, averaged over a 10-second interval. In the cases when complete time series data is not available for a session, the charging power is assumed constant such that the total charging energy delivered is consistent with the primary data. The disaggregated session data is summed into a total site charging power time series, which is then averaged over 15-minute intervals. Timestamp indices refer to the beginning of the 15-minute interval. The data period is from 2018-5-1 00:00 to 2019-2-28 23:45.

%\hypertarget{solar-production}{%
\subsubsection{Solar Production}\label{solar-production}%}

The modeled PV production power begins life as Geostationary Operational Environmental Satellite (GOES) satellite solar irradiance data from the US National Solar Resource Database (NSRDB) \cite{Sengupta2018}. These data are specific to the years in question and are not from a typical meteorological year (TMY) set. See Table \ref{tab:data-summary}. Prism Solar 350 W bifacial solar module DC electrical power is estimated using the California Energy Commission Performance Model, a 6-parameter physical PV cell model. The full AC array power is estimated given loss assumptions and an inverter efficiency lookup table in the US National Renewable Energy Laboratory (NREL) System Advisor Model (SAM) database \cite{NREL2022} for the Enphase IQH 380 W microinverter. All these functions are implemented in SAM v2022.11.21 \cite{NREL2022}. Five different solar array orientations are modeled and simulated in different case studies, summarized in Table \ref{tab:casestudies}. Each orientation has the same number and type of bifacial modules. In the first orientation, all modules face South at a 20° tilt. In the remaining four orientations a percentage of the modules face all modules face West at 90° tilt: 25\%, 50\%, 75\%, and then 100\%. Tilt is defined as 90° minus the altitude angle of the normal vector of the primary active face of the module. No shading is considered. Of the two sides of the bi-facial module, the primary active face is oriented West since the afternoon load is generally subject to higher prices.
